### 2015-04-29 Doc/paper.tex
start on report

%\documentclass[conference]{IEEEtran}

\documentclass[runningheads,a4paper]{llncs}

\usepackage{graphicx}
\usepackage[vlined,ruled]{algorithm2e}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{listings-golang}
\usepackage{hyperref}

\definecolor{lightgray}{rgb}{.9,.9,.9}
\definecolor{darkgray}{rgb}{.4,.4,.4}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mygreen}{rgb}{0,0.6,0}

\hypersetup{
    colorlinks,
    linkcolor={red},
    citecolor={blue},
    urlcolor={blue}
}

\lstset{ % add your own preferences
    basicstyle=\footnotesize,
    keywordstyle=\color{red},
    numbers=left,
    numbersep=5pt,
    showstringspaces=false, 
    stringstyle=\color{blue},
    tabsize=4,
    language=Golang % this is it !
}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Inferring Invariants of Distributed Programs}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{Arash Vahabzadeh \and Keheliya Gallaba}
\institute{University of British Columbia,\\
Vancouver, Canada\\
\mailsa}

\urldef{\mailsa}\path|{arashvhb,kgallaba}@ece.ubc.ca|

% \author{\IEEEauthorblockN{Keheliya Gallaba\IEEEauthorrefmark{1},
% Arash Vahabzadeh\IEEEauthorrefmark{2}}
% \IEEEauthorblockA{University of British Columbia\\
% Vancouver, BC, Canada\\
% Email: \IEEEauthorrefmark{1}kgallaba@ece.ubc.ca,
% \IEEEauthorrefmark{2}arashvhb@ece.ubc.ca}}

% make the title area
\maketitle

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

\input{introback}

\input{approach}

\input{evalrelate}

\bibliographystyle{IEEEtran}
\bibliography{paper}

% that's all folks
\end{document}